Global adjoint tomography -- Model GLAD-M25

### 2018-10-20 main.tex
Updates from ShareLaTeX

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the Northern Hemisphere and poor coverage in the Southern Hemisphere.
To balance the coverage,
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}

{\color{Red} Wenjie: is it $\omega_{cr}^{-2}$ or $\omega_{cr}^{-1}$? And does the normalization factor~$N_c$ depend on~$c$?\\ }
{\color{Red} To Jeroen: I checked the code. we are using $\omega_{cr}^{-1}$. The normalization factor depends on c.}

where~$N_c$ is a normalization factor,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a spatial distribution of
points on the unit sphere, determine a weighting associated with each point.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/events_1480.pdf}
  \caption{1,480 earthquakes used in this study. (a) Distribution of earthquakes. The color of each beach ball reflects its depth range, where blue designate events shallower than 50~km, green events between 50~km and 300~km, and red events deeper than 300~km. (b) and (c) Histograms of earthquake moment magnitudes and depths.}
  \label{fig:event_1480}
\end{figure}

@@

{\color{red} Wenjie: is there just one $\Delta$, or are there different ones for sources, receivers and categories? And is there just one normalization constant~$N$, or are there multiple ones?}

{\color{red} To Jeroen: You are right. $\Delta$ is different for a given source and category. Actually, we don't directly control $\Delta$. Instead, we control the condition number of weights. So the $\Delta$ will vary with receiver distributions. }

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red diamond) waves and surface waves
  (red square).
  {\color{red} Wenjie, why do the graphs overlap at iteration 22, but not at iteration 18? Shouldn't 22 look like 18, i.e., no overlapping points?    \textbf{To Jeroen}: overlap means I calculate the misfit twice using the same mode. For example, in iteration M22, I caculated the misfit using the 1040 earthquakes and 1480 earthquakes, respectively. In M18, I haven't done so. If you want, I can calculate the misfit values and adding the overlap misfit point.}
  }
  \label{fig:misfit}
\end{figure}

@@

{\color{red} Wenjie: Make sure the variance reductions are calculated as ln (Phi\_M25 / Phi\_M15).}
{\color{red} Wenjie: I was using (Phi\_M25 - Phi\_M15) / Phi\_M15. So let me modify it?}
\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical (\%) & Radial (\%) &  Transverse (\%) \\
    \hline
    17--40~s  body waves    &   17.4 (31.1)   &       21.1 (37.1) &       24.6 (42.2) \\
    40--100~s body waves    &   16.8 (31.3)  &       21.3 (40.2)  &       21.7 (42.0) \\
    40--100~s surface waves &   28.5 (50.1)  &       28.3 (51.1) &       28.4 (53.8)  \\
    90--250~s surface waves &   14.5 (36.2)  &       13.4 (38.8) &       25.7 (38.6) \\
    \hline
  \end{tabular}\\
  \caption{Changes in fit in the twelve measurement categories
between the new model, GLAD-M25, and its starting model, GLAD-M15, and, in parentheses, between GLAD-M25 and model S362ANI combined with CRUST2.0, which was the starting model for the GLAD-M15 inversion.}
  \label{table:misfit_reduction_M15_M25}
\end{table}

@@

First, we repeated the misfit assessment discussed in Section~\ref{section:Misfit assessment} for the held-out database.
Table~\ref{table:misfit_reduction_M15_M25_360} summarizes the reductions in misfit for the held-out database for model GLAD-M25 compared to GLAD-M15 as well a S362ANI combined with CRUST2.0.
It is comforting to see that the reductions are comparable to the
values tabulated in Table~\ref{table:misfit_reduction_M15_M25} for the actual inversion.
Overall the reductions are slightly smaller, partly due to the fact that we did not conduct full CMT inversions for the held-out database.

@@

Second, we repeated the histogram comparisons discussed in Section~\ref{section:Histogram comparisons} for the held-out database.
We used a set of identical windows on all three components to assess
phase and amplitude anomalies in models GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
Fig.~\ref{fig:phase_hist_360} shows histograms of the resulting phase
measurements in all twelve measurement categories.
The total number of measurements exceeds 7.8~million.
As in the actual inversion,
we observe systematic decreases in all standard deviations in all twelve categories,
and we note a sharpening of the distributions centered around zero.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/dt_histogram_360.pdf}
  \caption{Histogram of phase measurements for the 360 events in the held-out database.}
  \label{fig:phase_hist_360}
\end{figure}

@@

We first examine model GLAD-M25 in the context of existing global tomographic models.
In Fig.~\ref{fig:global-vs} we compare global maps of the isotropic part of our shear wavespeed model with model S362ANI$+$M~\citep{moulik2014anisotropic},
which is an updated version of the starting model for the GLAD-M15 inversion,
and model S40RTS~\citep{ritsema2011s40rts}.
The range of perturbations in shear wavespeed varies from map to map, as indicated in the top left of each panel.
Overall these models are in good agreement at the longest wavelengths, especially Glad-M25 and S362ANI$+$M.
The perturbations in GLAD-M15 tend to be larger than in the other models,
consistent with observations by~\cite{french2014whole},
whose model is also based on a form of waveform inversion.

@@

An important aspect of our inversion is that constrain shear and compressional waves simultaneously.
In Fig.~\ref{fig:global-vs} we compare global maps of our compressional wavespeed model with P models LLNL-G3Dv3~\citep{simmons2012llnl} and GAP-P4~\citep{fukao2013subducted}.
At a depth of 100~km,
GLAD-M25 shows prominent mid-oceanic ridges, cooling of the Pacific lithosphere, and the roots of numerous continental cratons.
In the upper mantle, LLNL-G3Dv3 exhibits the largest lateral variations,
e.g., very slow compressional wavespeeds at 500~km depth below Hawaii, Samoa, and Tahiti.
In the midmantle we can see remnants of subduction beneath South America, Indonesia,
and Tonga-Kermadec.
In the deep mantle, LLNL-G3Dv3 and GAP-P4 begin to fade out, but the long-wavelength patterns,
in particular the lowerparts of the Africa and Pacific superplumes, are very similar.

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vp_LLNL-GAP.pdf}
  \caption{Map views of global compressional wavespeed variations at various depths for model
  GLAD-M25 (left column), LLNL-G3Dv3 \citep[middle column;][]{simmons2012llnl} and
  GAP-P4~[right column;][](\citep{fukao2013subducted}).
  For LLNL-G3Dv3, depth ranges are labeled on the bottom right
  bottom of each panel. Other plotting conventions are as in Fig.~\ref{fig:global-vs}.}
\label{fig:global-vp}
\centering
\end{figure}

@@

The deployment of USArray has dramatically increased data overage underneath the
continental United States.
In this section, we compare GLAD-M25 with
with two global models, S362ANI$+$M and S40RTS, and two regional models,
US22~\citep{zhu2017radial} and SL2013NA~\citep{schaeffer2014imaging}.
US22 is a radially anisotropic model based on adjoint tomography using
USArray data from 180 regional earthquakes using 15--50~s
body waves and 25--150~s surface waves.
SL2013NA is a global upper-mantle shear wavespeed model based on multimode
surface waveform tomography primarily focused on North America.

North America is characterized by a large, high-wavespeed cratonic
lithosphere, bounded by the Rocky Mountain Front to the West and continental
margin to the East. On both sides the lithosphere has been deformed by
low wavespeed structures, thus forming sharp boundaries.

At 100km, our model shows a much sharpened boundary between the carton
and surrounding low velocity structures, especially compared with
two global models, S362ANI$+$M and S40RTS, yet stay very similar to
two regional models, US22 and SL2014NA. At such depth,
we are observed a very strong expression of the Snake River Plain(
with Yellowstone) and Raton hotspot, with their depth extending to 200km.
The very high-velocity craton extend to 200km, with its very core part to
300 km, similar to US22 while SL2013NA shows it almost disappeared at such depth.
Again, the Wyoming craton and Medicine Hat Block shows a very strong lithosphere
root from the very shallow depth to 400km, which is also similar to US22 but
hard to observe in SL2013NA. Similar case is the craton at Yavapai and
Mazatzal Blocks, which define the southern bound of the large craton.

At such depth, we also observed high velocity
features within the Gulf of Mexico, which coincides spatially with the deepest
bathymetry and corresponds to a portion of ancient oceanic
Lithosphere\citep{muller2008}. The remnant of Juan de Fuca subduction
plates, with its high velocity footprint, is presented in our model at
300km and 400km, though the SL2013NA shows a much stronger anomaly.

At the North, our models shows a extended high velocity craton beneath the 
Greenland, though clearly separated with the North America Craton,
bounded at the relatively low velocity structure at Barffin Bay and Labrador
Sea, with North American Craton followed nicely with the coastline. In the
north west region, our models shows the most clear picture of
Aleutian subduction zones, with it high velocity feature down to 400km.
In the northeastern part, the Iceland hotspot has been greatly enhanced,
especially at 300km and 400km compared to S362ANI$+$M.

In the Easter part, our models shows a high-velocity structure at the western
Atlantic, starting from the tip of Florida to the New York,
separated from the major part the north America craton,
though it stays very thin that the depth only extend to above 200km.
This feature is associated with some of the oldest oceanic crust and lithosphere
within the Atlantic\citep{muller2008, schaeffer2014imaging}.
All the models is showing a high velocity anomaly at Nova Scotia and Newfoundland
with its high-velocity core to 300km.
Bermuda hotspot is nicely preserved from our starting
model, yet get sharpened and stronger at 200km and 300km. Besides, the
Caribbean suduction, which is relatively young subduction systems, is clearly
observable in our model, starting from 200km to 300km.

For even deeper depth, our models shows a very clear picture of the Farallon
subductions, but it goes beyond the range of this comparison since the two
region models are really upper-mantle models which don't extend to the very
deep depth. We will show some vertical cross-sections of Farallon slabs in
the later sections.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/depth_slice/america_vs.pdf}
  \caption{Map views of $V_s$ variations of North America at various depths
  for our model GLAD-M25(first column) and several global(S362ANI$+$M and S40RTS)
  and regional(US22\citep{zhu2017radial} and SL2013NA\citep{schaeffer2014imaging})
  models. Green circles denotes the locations of hotspots. For SL2013NA, we
  showed the 610km instead of 600km due to its own mesh.}
\label{fig:america-vs}

@@

Fig.\ref{fig:europe-vs} shows the isotropic $V_s$ structures at Europe. As
comparisons, we also showed S362ANI$+$M, S40RTS and one regional model EU60
\citep{zhu2015seismic}. Compared to two global models, it is clear that
our model is much sharper at few locations. At 100 and 200km, Eifel
hotspot is clearly seen in our models, as well as in EU60. At 400km
and 600km, the subduction is very consistent between GLAD-M25 and EU60,
with the curved shape at the lower bottom of the figure,
while S362ANI$+$M shows very smooth shape and S40RTS shows much smaller 
perturbation values. Besides, at 400km, GLAD-M25 is enhancing Iceland
compared to S362ANI$+$M.

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/europe_vs.pdf}
  \caption{Map views of $V_s$ variations of Europe at various depths for our
  model GLAD-M25(first column) and global(S362ANI$+$M and S40RTS) and
  regional(EU60\citep{zhu2015seismic}) models. For EU60, the range of color
  bar is 4\% for all depths, which is labeled in the right bottom of the figures.}
\label{fig:europe-vs}
\centering
\end{figure}

@@

Asia has a very complicated tectonic settings, containing
the interaction among Indian, Euroasian, Australia, Pacific and Philippine
plates. Even though we don't have a significant large number of high
event-response stations at this region(compared to Europe and North
America), our inversion is compensated by the large number of temporary
arrays and earthquakes.

Our model at Asian region is plotted on Fig.\ref{fig:asia-vs},
together with S362ANI$+$M, S40RTS and EARA2014\cite{chen2015multiparameter}.
EARA2014 is a regional 3D anisotropic tomographic model using adjoint
methods, comprising 1.7million frequency-dependent traveltime
measurements from 227 earthquakes recorded by 1869 seismic stations.
It has superior data coverage, mainly by incorporating the China
Seismic Array in this area.

The southwestern part of China is featured by the Himalayas mountain,
which is 
a shallow expression of the collision between India and Euroasian plate. Our
models shows a strong high-velocity anomalies near the plate boundary, and this
anomaly has all the way extending from 100km to 600km depth. With the increase
of depth, the high-velocity regions get sharper, indicating the narrowing
root of subduction plate. The high-elevation Tibetan plateau, sitting to
the north of Himalaya mountains, shows a relatively low velocity
comparing to its surroundings, including the Himalayas at its south and
Tianshan Mountain at is North, which is consistent with other global
and regional studies.

The southern part of Asia sees the majority tectonic activity between the Philippine
, EuroAsian and Australia plate. The Sumatra and Java trench, shown as sharp high-velocity
anomaly in our models below 200km, nicely follows the boundary between Euroasian
and Australia plates. This feature stays still very sharp at the 600km and extend
down to 1000km. We will have more detailed discussion about it in the following sections.
The Malina and Philippine Trench is also much sharper in our model between 200km and
600km.
The eastern part has a series of trenches, including Japan Trench, Izu-Bonin Trench,
Mariana Trench from North to the South. All those features follows perfectly with
the plate boundary, as emerging features compared the two global models from 200km
to 400km depth. At 100km depth, our model is strengthening the low-velocity structures
in the east of Philippine Plate and the anomaly gradually moved to the west at 200km
depth. In the north, our models shows low-velocity structures associated
with the Altay-Sayan and Baikai rift systems.

Other than the boundaries, our models also show much more details at the
intra-plate tectonic activities. It has good correlation of low-velocity regions with
the Hainan spot and Changbai Volcanoes. Those low-velocity features is prominent at
shallow depth, but gradually reduce its magnitude down with the increase of
depth, indicating that those features might be lithosphere's activity rather
than deep mantle's.
The Sichuan basin, a very localized high-velocity structures, is extending below
250km depth. At 600km, most of the models shows a large regions of high-velocities
near the Philippine plate, indicating the pounding of ancient subducted plates.

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/asia_vs.pdf}
  \caption{Map views of $V_s$ variations of Asia at various depths for our model GLAD-M25(first column) and several global(S362ANI$+$M and S40RTS) and regional(EARA2014\citep{chen2015multiparameter}) models.}
\label{fig:asia-vs}
\centering
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/south_america_vp.pdf}
  \caption{Map views of $V_p$ variations of South America at various depths for our model GLAD-M25(first column) and several global models, including S362ANI$+$M, LLNL-G3Dv3 and GAP-P4\citep{fukao2013subducted}.}
\label{fig:southamerica-vp}
\centering
\end{figure}

\subsection{Plumes}

@@

\subsection{Subductions}

>>>>

To determine the source parameters in the starting model,
we used the CMT inversion algorithm of \cite{liu2004spectral}.
This algorithm combines a normalized waveform difference misfit with an envelope difference misfit.
Specifically, the algorithm minimizes the misfit function
\begin{equation}
   \begin{split}
      \Phi =  \sum\limits_{c=1}^{C} \omega_c \sum\limits_{r=1}^{R_c} \omega_{cr}
       \sum\limits_{w=1}^{W_{cr}} \omega_{crw}\,
          & \left\{ \lambda\, \frac
              { \int \big[ d_w(t) - s_w(t - \Delta t) \big]^2 \mathrm{d}t}
              {\int \big[ d_w(t) \big]^2  \mathrm{d}t} \right.
       \\ & \quad \left. \mbox{} + (1 - \lambda)\, \frac
              {\int \big[ e(d_w(t)) - e(s_w(t - \Delta t)) \big]^2 \mathrm{d}t}
              {\int \big[ e(d_w(t)) \big]^2\mathrm{d}t} \right\}
              \quad ,
   \end{split}
\end{equation}
where~$d_w(t)$ denotes data in time window~$w$\,,
and~$s_w(t - \Delta t)$ the corresponding synthetic for model~GLAD-M15.
We allow the synthetics to shift relative to the data by an amount~$\Delta t$
--- effectively a station correction --- determined by cross correlation.
The envelope function is denoted by~$e(\,\cdot\,)$\,,
and the parameter~$\lambda$ determines the balance between fitting waveforms versus fitting envelopes.
The number of measurement categories is denoted by~$C$\,.
Following~\cite{ekstrom2012global},
seismograms were filtered between 50~s and 100~s
to select three-component body-wave windows,
and between 60~s and 100~s to select three-component
surface-wave windows,
resulting in six measurement categories, $c=1,\ldots,C$\,.
To balance their contributions,
each category is weighted by the reciprocal of the number of measurements in that
category, $\omega_c$\,.
The number of windows for a given receiver~$r$ in a given measurement category~$c$ is~$W_{cr}$\,.
Each such window is weighted equally, i.e.,~$\omega_{crw}=1$\,.
The number of receivers that records data in category~$c$ is denoted by~$R_c$\,.
As shown in Fig.~\ref{fig:source_correction}b,
the receivers are unevenly distributed across the globe,
with several dense arrays in the Northern Hemisphere and poor coverage in the Southern Hemisphere.
To balance the coverage,
we assign receiver weights~$\omega_{cr}$ based on the expression
\begin{equation}
\omega_{cr}^{-1} = N_c\,\sum_{r'=1}^{R_c} \exp\left[\mbox{}-\left(\frac{\Delta_{rr'}}{\Delta_c}\right)^2\right]
\quad ,
\label{eq:spatial_weights}
\end{equation}
where~$N_c$ is a normalization factor for category~$c$,
and where~$\Delta_{rr'}$ denotes the angular distance between receivers~$r$ and~$r'$\,.
The reference angular distance~$\Delta_c$ for each category~$c$ needs to be chosen such that the condition 
number of the diagonal weighting matrix defined by Eqn.~(\ref{eq:spatial_weights}) is not too large.
The calculation of the weights may be abstracted as: given a spatial distribution of
points on the unit sphere, determine a weighting associated with each point.

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/events_1480.pdf}
  \caption{1,480 earthquakes used in this study. (a) Distribution of earthquakes. The color of each beach ball reflects its depth range, where blue designate events shallower than 50~km, green events between 50~km and 300~km, and red events deeper than 300~km. (b) \& (c) Histograms of earthquake moment magnitudes and depths.}
  \label{fig:event_1480}
\end{figure}

@@

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/misfit.pdf}
  \caption{Evolution of phase misfits from GLAD-M15 to GLAD-M25.
  Each color denotes the different stages of inversions, between
  which we make changes to the inversion by either tuning the
  weighting strategies or adding more data.
  (a) Evolution of the total phase misfit. (b) - (j) Misfit in
  each period band and component. (e) - (g) From GLAD-M18,
  we future split the 40-100 sec period band into two
  measurement categories: body(red diamond) waves and surface waves
  (red square).
  {\color{red} Wenjie, why do the graphs overlap at iteration 22, but not at iteration 18? Shouldn't 22 look like 18, i.e., no overlapping points?    \textbf{To Jeroen}: overlap means I calculate the misfit twice using the same mode. For example, in iteration M22, I caculated the misfit using the 1040 earthquakes and 1480 earthquakes, respectively. In M18, I haven't done so. If you want, I can calculate the misfit values and adding the overlap misfit point.}
  {\color{blue}Wenjie: OK, then let's be consistent in what we do. I would prefer no overlap.}
  }
  \label{fig:misfit}
\end{figure}

@@

{\color{red} Wenjie: Make sure the variance reductions are calculated as ln (Phi\_M25 / Phi\_M15).}
{\color{red} Wenjie: I was using (Phi\_M25 - Phi\_M15) / Phi\_M15. So let me modify it?}
{\color{blue} Wenjie: yes please. It will only change things slightly, if at all. You need to do the same for Table~\ref{table:misfit_reduction_M15_M25_360}.}
\begin{table}[!htb]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    ~          &  Vertical (\%) & Radial (\%) &  Transverse (\%) \\
    \hline
    17--40~s  body waves    &   17.4 (31.1)   &       21.1 (37.1) &       24.6 (42.2) \\
    40--100~s body waves    &   16.8 (31.3)  &       21.3 (40.2)  &       21.7 (42.0) \\
    40--100~s surface waves &   28.5 (50.1)  &       28.3 (51.1) &       28.4 (53.8)  \\
    90--250~s surface waves &   14.5 (36.2)  &       13.4 (38.8) &       25.7 (38.6) \\
    \hline
  \end{tabular}\\
  \caption{Changes in fit in the twelve measurement categories
between the new model, GLAD-M25, and its starting model, GLAD-M15, and, in parentheses, between GLAD-M25 and model S362ANI combined with CRUST2.0, which was the starting model for the GLAD-M15 inversion.}
  \label{table:misfit_reduction_M15_M25}
\end{table}

@@

First, we repeated the misfit assessment discussed in Section~\ref{section:Misfit assessment} for the held-out database.
Table~\ref{table:misfit_reduction_M15_M25_360} summarizes the reductions in misfit for the held-out database for model GLAD-M25 compared to GLAD-M15 as well a S362ANI combined with CRUST2.0.
It is comforting to see that the reductions are comparable to the
values tabulated in Table~\ref{table:misfit_reduction_M15_M25} for the actual inversion.
Overall, the reductions are slightly smaller, partly due to the fact that we did not conduct full CMT inversions for the held-out database.

@@

Second, we repeated the histogram comparisons discussed in Section~\ref{section:Histogram comparisons} for the held-out database.
We used a set of identical windows on all three components to assess
phase anomalies in models GLAD-M25, GLAD-M15, and S362ANI combined with CRUST2.0.
Fig.~\ref{fig:phase_hist_360} shows histograms of the resulting phase
measurements in all twelve measurement categories.
The total number of measurements exceeds 7.8~million.
As in the actual inversion,
we observe systematic decreases in all standard deviations in all twelve categories,
and we note a sharpening of the distributions centered around zero.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/dt_histogram_360.pdf}
  \caption{Histograms of phase measurements for the 360 events in the held-out database.}
  \label{fig:phase_hist_360}
\end{figure}

@@

We first examine model GLAD-M25 in the context of existing global tomographic models.
In Fig.~\ref{fig:global-vs} we compare global maps of the isotropic part of our shear wavespeed model with model S362ANI$+$M~\citep{moulik2014anisotropic},
which is an updated version of the starting model for the GLAD-M15 inversion,
and model S40RTS~\citep{ritsema2011s40rts}.
The range of perturbations in shear wavespeed varies from map to map, as indicated in the top left of each panel.
Overall, these models are in good agreement at the longest wavelengths, especially Glad-M25 and S362ANI$+$M.
The perturbations in GLAD-M15 tend to be larger than in the other models,
consistent with observations by~\cite{french2014whole},
whose model is also based on a form of waveform inversion.

@@

An important aspect of our inversion is that it constrains shear and compressional waves simultaneously.
In Fig.~\ref{fig:global-vs} we compare global maps of our compressional wavespeed model with P models LLNL-G3Dv3~\citep{simmons2012llnl} and GAP-P4~\citep{fukao2013subducted}.
At a depth of 100~km,
GLAD-M25 shows prominent mid-oceanic ridges, cooling of the Pacific lithosphere, and the roots of numerous continental cratons.
In the upper mantle, LLNL-G3Dv3 exhibits the largest lateral variations,
e.g., very slow compressional wavespeeds at 500~km depth below Hawaii, Samoa, and Tahiti.
In the midmantle we can see remnants of subduction beneath South America, Indonesia,
and Tonga-Kermadec.
In the deep mantle, LLNL-G3Dv3 and GAP-P4 begin to fade out, but the long-wavelength patterns,
in particular the lowerparts of the Africa and Pacific superplumes, are very similar.

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/globe_vp_LLNL-GAP.pdf}
  \caption{Map views of global compressional wavespeed variations at various depths for model
  GLAD-M25 (left column), LLNL-G3Dv3 \citep[middle column;][]{simmons2012llnl} and
  GAP-P4~\citep[right column;][]{fukao2013subducted}.
  For LLNL-G3Dv3, depth ranges are labeled in the bottom right
  bottom of each panel. Other plotting conventions are as in Fig.~\ref{fig:global-vs}.}
\label{fig:global-vp}
\centering
\end{figure}

@@

The deployment of USArray has dramatically increased data overage across the
continental United States.
In this section, we compare GLAD-M25 in this region with
with two global models, S362ANI$+$M and S40RTS, and two regional models,
US22~\citep{zhu2017radial} and SL2013NA~\citep{schaeffer2014imaging}.
US22 is a radially anisotropic model based on adjoint tomography using
USArray data from 180 regional earthquakes using 15--50~s
body waves and 25--150~s surface waves.
SL2013NA is a global upper-mantle shear wavespeed model based on multimode
surface waveform tomography primarily focused on North America.

North America is characterized by a large, high-wavespeed lithospheric craton,
bounded by the Rocky Mountain Front to the West and continental
margin to the East. On both sides the lithosphere has been deformed by
low wavespeed structures, thus forming sharp tectonic boundaries.

At a depth of 100~km, we observed that the edges of the craton are sharper in
GLAD-M15 than in the other two global models,
and quite similar in shape and sharpness to the two regional models.
We can clearly see impressions of the Snake River Plain (with Yellowstone)
and the Raton hotspot, which also have distinct slow expressions at depths of 200~km.
The core of the craton exhibits fast wavespeeds at depths reaching 300~km Northwest of
Hudson Bay.
In the regional models the craton has basically faded away at these depths.
The Wyoming and Medicine Hat Blocks show a very strong and thick lithospheric
root from shallow depths to $\sim$400~km, similar to US22 but
difficult to discern in SL2013NA.
Similar observations can be made for the Yavapai and
Mazatzal Blocks, which define the southern boundary of the large craton.

In the 200--400~km depth range, we observe a high wavespeed
anomaly within the Gulf of Mexico, which coincides spatially with the deepest
bathymetry and corresponds to a portion of ancient oceanic
Lithosphere~\citep{muller2008}.
A remnant of the Juan de Fuca plate,
with a high wavespeed footprint, is visible at depths of 300~km and 400~km,
where SL2013NA shows a much stronger anomaly.

In the North,
we see a high wavespeed craton beneath Greenland which is partly separated from the
North American craton by low wavespeed structures beneath Baffin Bay and the Labrador Sea,
where the North American craton nicely follows the coastline.
In the Northeast corner of the maps we see that the Iceland hotspot is very prominent in GLAD-M25 in all depth ranges, similar to S40RTS but largely missing in S362ANI$+$M.
In the Northwest we see a clear imprint of the Aleutian subduction zone in the 200--400~km depth range, comparable to SL2013NA bu absent in the other global models.

In the East, at 100~km depth,
we see ageing of the Atlantic lithosphere, comparable to the other global models~\citep{muller2008, schaeffer2014imaging}.
All models show fast anomalies beneath Newfoundland and Nova Scotia extending to depths of 300~km.
The Bermuda hotspot shows a very nice slow wavespeed impression,
getting sharper and stronger at depths of 200~km and 300~km (note differences in the scales of the maps).
The Caribbean slab, a relatively young subduction zone, is clearly
visible in our model in the 200--400~km depth range.
At greater depths,
GLAD-M25 shows a very distinct image of the Farallon slab,
but since it goes well beyond the depth range of this comparison,
we return to this feature in Section~\ref{section:slabs}.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{figures/depth_slice/america_vs.pdf}
  \caption{Map views of isotropic shear wavespeed variations beneath North America 
  for model GLAD-M25 (first column) and several global (S362ANI$+$M and S40RTS)
  and regional~\citep[US22;][]{zhu2017radial} and~\citep[SL2013NA;][]{schaeffer2014imaging}
  models. Green circles denote the locations of hotspots. For SL2013NA, we
  show a depth of 610~km instead of 600~km, reflecting its model parametrization.}
\label{fig:america-vs}

@@

Fig.~\ref{fig:europe-vs} shows isotropic shear wavespeed perturbations beneath Europe.
For comparison,
we again include the two global models S362ANI$+$M and S40RTS,
as well as regional model EU60~\citep{zhu2015seismic},
which is based on adjoint tomography.
Again there is good agreemment with the two global models with regards to the longer wavelength structure,
although the Iceland hotspot is less prominent in S362ANI$+$M.
In the 100--200~km depth range we see the Massif Central, the Eiffel hotspot,
and the Central Slovakia Volcanic Field as distinctly slow anomalies,
in agreement with EU60.
In the 400--600~km depth range,
GLAD-M25 is again in good agreement with~EU60,
featuring past subduction in the Mediterranean,
including the present-day Aegean slab.

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/europe_vs.pdf}
  \caption{Map views of isotropic shear wavespeed variations beneath Europe for   
  model GLAD-M25 (first column), two global model (S362ANI$+$M and S40RTS), and
  one regional model~\citep[EU60;][]{zhu2015seismic}. For EU60, the range of the color
  bar is 4\% at all depths, as indicated in the bottom right bottom of the panels.}
\label{fig:europe-vs}
\centering
\end{figure}

@@

Asia, like Europe, has a complicated tectonic history, involving
interactions among the Indian, Eurasian, Australian, Pacific, and Philippine
plates.
Map views of GLAD-M15 centered on Asia are plotted in Fig.~\ref{fig:asia-vs},
together with S362ANI$+$M, S40RTS, and regional model EARA2014~\citep{chen2015multiparameter}.
EARA2014 is a transversely isotropic model based on adjoint
tomography, assimilating 1.7~million frequency-dependent traveltime
measurements from 227 earthquakes recorded by 1,869 seismic stations.
It has superior data coverage, mainly by incorporating data from the China Array.

The Southwestern part of China is dominated by the Himalayas,
expressing the collision between India and Eurasia.
GLAD-M25 exhibits strong high wavespeed anomalies in this area, which extend all the way from 100~km to 600~km, narrowing and sharpening with depth,
in good agreement with EARA2014.
The Tibetan plateau shows relatively low wavespeeds
compared to its surroundings, including the Himalayas to its South and
the Tianshan Mountains to its North, consistent with the other global
and regional studies.

The southern part of Asia shows major tectonic activity between the
Philippine, Eurasian, and Australia plates.
The Sunda trench is visible as a distinct high wavespeed anomaly below 200~km,
nicely following the plate boundary.
This feature remains very sharp at 600~km and extends
down to 1000~km.
We discuss this subducting slab further in Section~\ref{section:slabs}.
The Manila and Philippine trenches are also much sharper in our model between 200~km and
600~km than in the other global models, and comparable to EARA2014.
The eastern edge of the map features a series of trenches following plate boundaries,
including Japan, Izu-Bonin, and Mariana.
These slabs are absent in the other global model,
but feature more prominently in EARA2014.
Along the northern edge we see low wavespeeds associated with the
Altai-Sayan and Baikal rift systems.
In terms of intra-plate activities,
at shallow depths we see clear low wavespeed impressions of the Hainan and Changbai volcanoes,
suggesting a lithospheric rather than deeper mantle origin.
The Sichuan basin, a very localized high-wavespeed structure, extends below
250~km depth. At 600~km, most models shows a large region of high wavespeed
beneath the Philippine plate, indicating ponding of ancient subducted plates.

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/asia_vs.pdf}
  \caption{Map views of shear wavespeed variations beneath Asia for GLAD-M25~(first column),
  global models S362ANI$+$M (second column) and S40RTS (third column), and regional model EARA2014~\citep[last column;][]{chen2015multiparameter}.}
\label{fig:asia-vs}
\centering
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.9\textwidth]{figures/depth_slice/south_america_vp.pdf}
  \caption{Map views of compressional wavespeed variations beneath South America for GLAD-M25~(first column) and global models S362ANI$+$M (second column), LLNL-G3Dv3~(third column), and GAP-P4~\citep[last column;][]{fukao2013subducted}.}
\label{fig:southamerica-vp}
\centering
\end{figure}

\subsection{Plumes}
\label{section:plumes}

@@

\subsection{Slabs}
\label{section:slabs}
